4-5-interactive

### 2022-04-27 4-1-basic.Rmd
fix merge conflict

Victoria government of Australia provides daily COVID case number by date, source, and local government area (LGA). This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics. The first five rows of both data are printed below:

@@

Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create). \code{as_cubble()} will automatically check the match of sites in both tables and emit a warning if a location has missing spatial or temporal information.

@@

Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from GHCN, starting from year `r range(weatherdata::historical_tmax$first_year)[2]`. \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. This example compares the maximum temperature in two periods: 1971 - 1975 and 2016 - 2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted by the station number: Australia GHCN station number starts with "ASN00" and is followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml). The second and third digits of BOM station number (7th and 8th in the GHCN number) define the state of the station, where 46 - 75 are New South Wales stations and 76 - 90 are Victoria stations. Filtering Victoria and New South Wales stations is an operation in the spatial dimension and hence uses the nested form:

@@

Filtering for the period 1971 ~ 1975 and 2016 ~ 2020 is an operation on the time dimension and the nested cubble needs to be switched to the long cubble by \code{stretch()}:

@@

A monthly average is used for both periods to smooth the maximum temperature and it is again an operation on the time dimension:

@@

A few stations do not have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported. These stations are filtered out by examining whether the summarised time series has 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

>>>>

The Victoria State Government in Australia provides daily COVID case numbers by date, source, and local government area (LGA). This data can be used to visualise COVID spread when combined with map information on LGA, available from the Australian Bureau of Statistics. The first five rows of both data are printed below:

@@

Cubble can be created with separate spatial and temporal tables in a list, with other arguments (\code{key}, \code{index}, and \code{coords}) introduced in section \@ref(create). \code{as_cubble()} will automatically check the matching of sites in both tables and emit a warning if a location has missing spatial or temporal information.

@@

The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for `r nrow(weatherdata::historical_tmax)` Australian stations from the GHCN starting from year `r range(weatherdata::historical_tmax$first_year)[2]`. \code{weatherdata::historical_tmax} is already in a cubble, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. This example compares the maximum temperature in two periods: 1971-1975 and 2016-2020 for stations in Victoria and New South Wales.

Stations in the two states can be subsetted by the station number: the Australian GHCN station numbers start with "ASN00" and are followed by the [Bureau of Meteorology (BOM) station number](http://www.bom.gov.au/climate/cdo/about/site-num.shtml). The second and third digits of the BOM station number (7th and 8th in the GHCN number) define the state of the station; the range 46-75 corresponds to New South Wales stations and the range 76-90 corresponds to Victorian stations. Filtering Victoria and New South Wales stations is a spatial operation and hence uses the nested form:

@@

Filtering for the time periods 1971-1975 and 2016-2020 is a time operation and the nested cubble needs to be switched to the long cubble form by \code{stretch()}:

@@

A monthly average is used for both periods to smooth the maximum temperature, which is another time operation:

@@

A few stations do not have records during the time period 1971-1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period are not reported. These stations are filtered out by examining whether the summarised time series has 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

### 2022-04-27 4-2-agg.Rmd
fix merge conflict

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the cluster with a glyph map. This example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from GHCN, records daily precipitation and maximum/ minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple kmean algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group each station belongs to. More complex algorithms can be used for other problems, as long as there is a mapping from each station to a cluster.

@@

To create a group level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:

@@

With the group level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

@@

Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data. Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number of printed in the middle of y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
knitr::include_graphics(here::here("figures/basic-agg.png"))
```

>>>>

In the previous example, there has already been some overlapping of the glyphs for a few stations near (151E, 34S) and (152E, 33S), which is a problem when mapping more stations at the national level. Aggregation can be helpful in grouping series into clusters before visualising the clusters with a glyph map. This example shows how to organise data at both levels with \code{switch_key()}.

\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for `r nrow(weatherdata::climate_full)` stations in Australia from 2016 to 2020. A simple $k$-means algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs. More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.

@@

To create a group-level cubble, use \code{switch_key()} with the new key variable, \code{cluster}:

@@

With the group-level cubble, \code{get_centroid()} is useful to compute the centroid of each cluster, which will be used as the major axis for the glyph map later:

@@

Long form cubble at both levels can be accessed through stretching the nested form and with access to both station and cluster-level cubbles, various plots can be made to understand the cluster. Figure \ref{fig:basic-agg} shows two example plots that can be made with this data. Subplot A is a glyph map made with the cluster level cubble in the long form and subplot B inspects the station membership of each cluster using the station level cubble in the nested form.

```{r basic-agg, out.width="100%", fig.cap="Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster."}
knitr::include_graphics(here::here("figures/basic-agg.png"))
```

### 2022-04-27 4-3-water-level.Rmd
fix merge conflict

Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges and this includes variables: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature. In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.

@@

From the map, a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with precipitation measured by the climate station. As introduced in Section 3.2, \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020. The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}. \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation. This puts precipitation \code{prcp}, as the independent. Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly To return all the pairs of the match, \code{temporal_min_match} can be set to 0.

@@

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series. There are four pairs of matches, all located in the middle of Victoria and the concurrent increase in precipitation and water level can be observed.

>>>>

The Bureau of Meteorology collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges. The collected variables include: electrical conductivity, turbidity, watercourse discharge, watercourse level, and water temperature. In particular, water level will interact with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather stations and water gauges in Victoria.

@@

From the map, one can see that a few water gauges and weather stations are close to each other and the fluctuation of the water level could be matched up with the precipitation measured by the nearby climate station. As introduced in Section \@ref(matching), \code{match_sites()} can be used to match one source of data with another source in a cubble. Here \code{Water_course_level} in \code{river} will be matched to \code{prcp} in \code{climate} in 2020. The two datasets need to be specified as the first two arguments and the variable to match can be specified in \code{temporal_by} using the \code{by} syntax in \code{join}. \code{temporal_independent} controls the variable used to construct the interval and the goal here is to see if the water level in the river will reflect precipitation. This puts precipitation \code{prcp}, as the independent variable. Given there is one year's worth of data, the number of peaks (\code{temporal_n_highest}) to consider is slightly raised from a default of 20 to 30 and \code{temporal_min_match} is raised accordingly. To return all the pairs of the match, \code{temporal_min_match} can be set to 0.

@@

Figure \ref{fig:matching} plots the matched pairs on the map. There are four pairs of matches, all located in the middle of Victoria and the expected concurrent increase in precipitation and water level can be observed.

### 2022-04-27 4-5-interactive.Rmd
fix merge conflict

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} object separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance. A ribbon band is constructed using the long form cubble to show each station's  maximum and minimum temperature across month. With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo-code that outlines the process to construct an interactive graphic described above:

@@

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum nad minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in Auguest in the seasonal display. A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))
#knitr::include_graphics(here::here("figures/linking-south.png"))
```

@@

Figure \ref{fig:interactive-popup} shows a reproduction of Figure \ref{fig:interactive-linking} made  with leaflet and popups [@leafpop].

>>>>

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016-2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} object separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance. A ribbon band is constructed using the long form cubble to show each station's maximum and minimum temperature across month. With a different dataset, users can calculate other per station measure in the nested form or make other time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo-code that outlines the process to construct an interactive graphic described above:

@@

```{r interactive-linking, echo = FALSE, out.width="100%", out.height="23%", fig.retina = 2, dpi = 300, fig.cap = "Exploring temperature variation using linking of a map and seasonal display. Each row is a screen dump of the process. The top row shows all locations and all temperature profiles. Selecting a location with high variance on the map produces the plot in the second row. The maximum and minimum temperature is shown using a ribbon. The bottom row first selects the lowest temperature in August in the seasonal display. A location in the Tasmania Island is then selected to compare the temperature variation with Thredbo AWS.", fig.show='hold'}
knitr::include_graphics(here::here("figures/linking.png"))
knitr::include_graphics(here::here("figures/linking-north.png"))
knitr::include_graphics(here::here("figures/linking-lower.png"))
#knitr::include_graphics(here::here("figures/linking-south.png"))
```

@@

Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made with leaflet and popups [@leafpop].
